Global patterns of forest autotrophic carbon fluxes

### 2020-03-02 manuscript/manuscript draft.Rmd
condensed first par of discussion

We found a significant relationship between length of growing season (temperature-defined) and FACF, with all fluxes showing a linear increase with length of growing season (Figs. 4e, S6-S7; Table S2). Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed.

@@

When FACF were standardized by growing season length, correlations with growing season climate--includingtemperature, precipitation, solar radiation, and PET--were generally weak (Table S4, Figs. S8-S9). Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similaryly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{woody-stem}$ ($R^2$ = 0.06, p<0.05). {**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.

@@

We used a large global database (ForC), containing an unprecedented amount of data representing all major forest types (Fig. 1) and the nine most significant forest autotrophic carbon fluxes (FACF) variables (Table 2), to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale. We test a suite of hypotheses, including many previously posed (Table 1), standardising for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.

bbl: Most of these two paragraphs (following) just restate the results. Honestly I would remove them enitrely, but at the very least, condense into a single paragraph.

We show that, across all nine FACF analyzed, C cycling decreases linearly with latitude (*H1.1*; Fig. 2)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**). The FACF increase in proportion to one another (*H1.2*), with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig. 2), indicating no major, systematic omissions or overestimations of flux components. However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_{woody}$*, *$BNPP_{fine.root}$*; Fig. 2; **some SI table?**).

**Climate explains a significant proportion of variation in all C fluxes analyzed** (albiet less for subsidiary fluxes), with temperature variables the best predictors of FACF at this global scale (Figs. 3-4). While other climate variables are significant predictors of FACF (*H3*), none of them in isolation improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig. 4). Water availability is an important factor in explaining FACF on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP (Fig. 4b) and a significant interaction between MAT and MAP for seven of the nine variables analyzed (Fig. 3). Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Fig. 4e-f), though growing season length doesn't improve upon MAT as a predictor. Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (*H5*; **some SI table?**). These findings clarify the big picture of how FACF vary with latitude and climate on a global scale.

>>>>

We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase with length of growing season (Figs. 4e, S6-S7; Table S2). Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed (Table S4).

@@

When FACF were standardized by growing season length, correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs. S8-S9). Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similaryly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{woody-stem}$ ($R^2$ = 0.06, p<0.05). {**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.

@@

Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale. We show that, across all nine FACF analyzed, C cycling decreases linearly with latitude (*H1.1*; Fig. 2)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**). The FACF generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig. S2) and with component fluxes summing appropriately to larger fluxes (Fig. 2), indicating no major, systematic omissions or overestimations of flux components. However, climate explained lower proportions of variability among subsidiary C fluxes (e.g., *$ANPP_{woody}$*, *$BNPP_{fine.root}$*, $R_{auto-root}$; Fig. 2; Table S2). Latitudianal variation in FACF is primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs. 3-4). Water availability is also influential, but generally of secondary importance across the range represented in our database (Figs. 3-4). Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Figs. 4e-f, S2, S6-S7), though growing season length doesn't improve upon MAT as a predictor. Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of FACF (*H5*; Fig. S9; Table S4). These findings clarify the big picture of how FACF vary with latitude and climate on a global scale.
